---
title: "Group Project - Clean Data"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Import Data Sets
This can take a while to run, these are relatively large .csv files
```{r}
movies = read.csv('../Data/movies_metadata.csv')
credits = read.csv('../Data/credits.csv')
```

#Describe movies data set
```{r}
colnames(movies)
```

Column definitions

|Field|Data Type|Definition|
|-----+----------------+-------------------------------------------------------------------------------------------------|
|adult|Boolean|Indicates if the movie is adult content (TRUE)|
|belongs_to_collection|List|If the movie is part of a collection, lists the name of the collection information|
|budget|Numeric|Dollar value of the production budget|
|genres|List|All genres associated with the film|
|homepage|String|Website of the film if available|
|id|Numeric|ID value for the film|
|imdb_id|String|IMDB URL Value for the film|
|original_language|String|Two letter code for production language|
|original_title|String|Working title of the film|
|overview|String|Brief description of the plot of the film|
|popularity|Numeric|Calculated popularity of the film|
|poster_path|String|URL for the film's poster|
|production_companies|List|All production companies involved in production of the film|
|production_countries|List|All countries involved in production of the film|
|release_date|Date|Date of film's theatrical release|
|revenue|Numeric|Revenue generated by the film|
|runtime|Numeric|Film run time in minutes|
|spoken_languages|List|All languages spoken in the film|
|status|String|Status of the films production. In Production, Cancelled, Post Production, Released etc.|
|tagline|String|Marketing tagline for the film if available|
|title|String|Title of the film when released|
|video|Boolean| If the film was released directly to video channels (non-Theatrical)|
|vote_average|Numeric|Average score on IMDB|
|vote_count|Numeric|Number of votes from IMDB users|

#Format and Clean Up Movies Data
```{r}
library(lubridate)
library(stringr)
library(tidyr)

movies.data = movies

#Convert Data Types As Necessary
movies.data$adult = as.logical(movies.data$adult)
movies.data$video = as.logical(movies.data$video)
movies.data$genres = as.character(movies.data$genres)
movies.data$release_date = ymd(movies.data$release_date)
movies.data$revenue = as.numeric(movies.data$revenue)
movies.data$budget = as.numeric(levels(movies.data$budget))[movies.data$budget]

#Transform Genres into Lists
movies.data$genres = gsub("\\[|\\]","",movies.data$genres)
movies.data$genres = gsub("\\{|\\}|'id': \\d+, 'name': |\\ \\{","",movies.data$genres)
movies.data$genres = gsub("'","",movies.data$genres)
movies.data = separate(movies.data, genres, into = paste("Genre", 1:10, sep='_'), sep=',')

#Transform List of Production Companies into fields
movies.data$production_companies = gsub("\\[|\\]","",movies.data$production_companies)
movies.data$production_companies = gsub("\\{'name': |, 'id': \\d+\\}| \\{'name': ","",movies.data$production_companies)
movies.data$production_companies = gsub("'","",movies.data$production_companies)
movies.data = separate(movies.data, production_companies, into = paste("Production_Company", 1:27, sep='_'), sep=',')

#Transform List of Production Countries into fields
movies.data$production_countries = gsub("\\[|\\]","",movies.data$production_countries)
movies.data$production_countries = gsub("\\{'iso_\\d+_\\d+': '\\D{2}', 'name': |\\}","",movies.data$production_countries)
movies.data$production_countries = gsub("'","",movies.data$production_countries)
movies.data = separate(movies.data, production_countries, into = paste("Production_Countries", 1:26, sep='_'), sep=',')

#Transform List of Spoken Languages into fields
movies.data$spoken_languages = gsub("\\[|\\]","",movies.data$spoken_languages)
movies.data$spoken_languages = gsub("\\{|\\}|'iso_\\d+_\\d+': '\\D{2}', 'name': ","",movies.data$spoken_languages)
movies.data$spoken_languages = gsub("'","",movies.data$spoken_languages)
movies.data = separate(movies.data, spoken_languages, into = paste("Spoken_Language", 1:20, sep='_'), sep=',')

#Transform Collection
movies.data$belongs_to_collection = str_extract_all(movies.data$belongs_to_collection,"(?<='name': ')(.*)(?=', 'poster_path)", simplify = TRUE)


#Add New Fields
movies.data$profit = movies.data$revenue - movies.data$budget

```
#Describe Credits Data Set
```{r}
colnames(credits)
```
|Field|Data Type|Definition|
|-----+----------------+-------------------------------------------------------------------------------------------------|
|cast|List|Every actor receiving billing in the credits, in order of IMDB Listing (Popularity)|
|crew|List|Every crew member receiving billing in the credits, in order of department and IMDB Listing (Popularity)|
|id|int|IMDB id for linking to other data|

#Format and Clean Up Crew Data
```{r}
library(dplyr)

credits.data = credits

#Top 10 billed Actors
credits.data$cast = gsub("\\[|\\]","",credits.data$cast)
credits.data$cast = str_extract_all(credits.data$cast,"(?<='name': ')(.*?)(?=', 'order':)",simplify = FALSE)
credits.data$cast = gsub("character\\(0|c\\(|\\)|\\\"","",credits.data$cast)
credits.data = separate(credits.data, cast, into = paste("ACTOR", 1:10, sep='_'), sep=',')

#Up to 3 Directors
credits.data$directors = str_extract_all(credits.data$crew,"(?<='Director', 'name': ')(.*?)(?=', 'profile_path':)",simplify = FALSE)
credits.data$directors = gsub("character\\(0|c\\(|\\)|\\\"","",credits.data$directors)
credits.data = separate(credits.data, directors, into = paste("Director", 1:3, sep='_'), sep=',')

#Up to 3 Screenplay
credits.data$screenplay = str_extract_all(credits.data$crew,"(?<='Screenplay', 'name': ')(.*?)(?=', 'profile_path':)",simplify = FALSE)
credits.data$screenplay = gsub("character\\(0|c\\(|\\)|\\\"","",credits.data$screenplay)
credits.data = separate(credits.data, screenplay, into = paste("Screenplay", 1:3, sep='_'), sep=',')

#Up to 3 Writers
credits.data$writer = str_extract_all(credits.data$crew,"(?<='Writer', 'name': ')(.*?)(?=', 'profile_path':)",simplify = FALSE)
credits.data$writer = gsub("character\\(0|c\\(|\\)|\\\"","",credits.data$writer)
credits.data = separate(credits.data, writer, into = paste("Writer", 1:3, sep='_'), sep=',')
#Convert Blank Cells to NA

empty_as_na <- function(x){
    if("factor" %in% class(x)) x <- as.character(x) ## since ifelse wont work with factors
    ifelse(as.character(x)!="", x, NA)
}

#Reformat Blanks
credits.data = credits.data %>% mutate_all(funs(empty_as_na)) 

```

```{r}
#Output data
write.csv(movies.data, '../Data/Processed Data/movies_data.csv')
write.csv(credits.data, '../Data/Processed Data/credits_data.csv')
```
